Conformance Checking Using Embeddings and the Applicability of Such Methods in the Internet of Things

### 2021-11-30 paper.tex
Extended some parts

This is done by keeping track of four numbers: \emph{\underline{p}roduced}, \emph{\underline{c}onsumed}, \emph{\underline{m}issing} and \emph{\underline{r}emaining} tokens during replay.
According to the semantics of petri nets, whenever a transition is fired it consumes one token from all input places and produces one token in each output place.
However since we are dealing with possibly non-conformant trace, it is possible that a trace contains activities, that are not enabled in the model at that point in time.
In this case a missing token is added to all input places, that did not contain a token, and the $m$ counter is incremented by the according amount.
Conversely, it might also happen, that a trace can be replayed, but that there are stray tokens left in the petri net after the replay. 
At the end of the replay the sum of all leftover tokens makes up the $r$ counter. \todo{Some explanation models could be included if the space is there for them}
Both missing and remaining tokens signal deviation from the process model, but both of these numbers are dependent on the length of traces.
The longer a trace is the more activities will be taken, the more tokens will be produced, and the more chances there are for tokens to be missing or remaining.
Importantly a trace where half the produced tokens were never consumed, i.e. remained, should intuitively have a lower fitness than one where only a quarter of produced tokens remained.
Of course the same relationship holds for the other two counters $m$ and $c$.
Then one very naturally arrives at a trace fitness defined as:
\begin{equation}
fitness(σ,N) = \frac{1}{2}(1-\frac{m_{N,σ}}{c_{N,σ}})+\frac{1}{2}(1-\frac{r_{N,σ}}{p_{N,σ}})
\end{equation}
, where $x_{N,σ}$ is the counter when replaying trace $σ$ on petri net $N$.
Often one does not just want to determine the fitness of one trace but a complete log, then this definition can be extended to 
\begin{equation}
	fitness(L,N) = \frac{1}{2}(1-\frac{\sum_{σ\in L}m_{N,σ}}{\sum_{σ\in L}c_{N,σ}})+\frac{1}{2}(1-\frac{\sum_{σ\in L}r_{N,σ}}{\sum_{σ\in L}p_{N,σ}})
\end{equation}
Note that log-fitness is not mathematically equivalent to the mean trace fitness.

@@

@@

>>>>

This is done by keeping track of four numbers: \emph{\underline{p}roduced}, \emph{\underline{c}onsumed}, \emph{\underline{m}issing} and \emph{\underline{r}emaining} tokens during replay.
According to the semantics of petri nets, whenever a transition is fired it consumes one token from all input places and produces one token in each output place.
However since we are dealing with possibly non-conformant trace, it is possible that a trace contains activities, that are not enabled in the model at that point in time.
In this case a missing token is added to all input places, that did not contain a token, and the $m$ counter is incremented by the according amount.
Conversely, it might also happen, that a trace can be replayed, but that there are stray tokens left in the petri net after the replay. 
At the end of the replay the sum of all leftover tokens makes up the $r$ counter. \todo{Some explanation models could be included if the space is there for them}
Both missing and remaining tokens signal deviation from the process model, but both of these numbers are dependent on the length of traces.
The longer a trace is the more activities will be taken, the more tokens will be produced, and the more chances there are for tokens to be missing or remaining.
Importantly a trace where half the produced tokens were never consumed, i.e. remained, should intuitively have a lower fitness than one where only a quarter of produced tokens remained.
Of course the same relationship holds for the other two counters $m$ and $c$.
Then one very naturally arrives at a trace fitness defined as:
\begin{equation}
fitness(\sigma,N) = \frac{1}{2}(1-\frac{m_{N,\sigma}}{c_{N,\sigma}})+\frac{1}{2}(1-\frac{r_{N,\sigma}}{p_{N,\sigma}})
\end{equation}
, where $x_{N,\sigma}$ is the counter when replaying trace $\sigma$ on petri net $N$.
Often one does not just want to determine the fitness of one trace but a complete log, then this definition can be extended to 
\begin{equation}
	fitness(L,N) = \frac{1}{2}(1-\frac{\sum_{\sigma\in L}m_{N,\sigma}}{\sum_{\sigma\in L}c_{N,\sigma}})+\frac{1}{2}(1-\frac{\sum_{\sigma\in L}r_{N,\sigma}}{\sum_{\sigma\in L}p_{N,\sigma}})
\end{equation}
Note that log-fitness is not mathematically equivalent to the mean trace fitness.

@@

\subsection{Embeddings}
Embeddings are a popular machine learning approach that has been used for a variety of tasks with the most successful applications being in the fields of natural language processing and computer vision.
The fundamental idea of embeddings is to train some form of embedding function that embeds complex objects, such as words or documents in the case of NLP or pictures and video sequences in the case of computer vision in a vectors space\footnote{To be more precise: a Hilbert space, since well-defined notions of distances and angles are necessary.}.
Such an embedding function then enables a proper mathematical treatment of these non-vector objects as vectors, which allows for the application of a broad range of algorithms.
Most importantly it makes it possible to define notions of distance and/or (dis-)similarity needed for machine learning and data analysis tasks.

\subsection{word2vec}

@@

\subsubsection{Trace embeddings}
In contrast to the two previous approaches, which used activity embeddings, it is also possible to train trace embeddings.
The relationship between \emph{act2vec} and \emph{trace2vec} is analogous to the one between \emph{word2vec} and \emph{doc2vec}.
This introduces additional instability on the trace level, since there will generally many more traces than activities, but allows for the direct application of distance metrics in the vector space when comparing two traces.
This is much less computationally expensive as the previous two approaches which were complex distance metric between two sets of embeddings.
